### 2020-04-29 PRD proofs/ProofCorrections_DN12236(PRD).tex
added proofs and response files

\documentclass[10pt,a4paper]{report}

%\usepackage{kpfonts}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage[french,english]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}

\usepackage{xcolor}

\usepackage[left=1cm,right=1cm,top=1cm,bottom=1.5cm]{geometry}

\usepackage{indentfirst}

\usepackage{academicons}
\definecolor{orcidlogocol}{HTML}{A6CE39}
\definecolor{darkbue}{rgb}{0.1,0,0.65}
\usepackage[pdftex,hidelinks]{hyperref}

%\usepackage{tikz}
%\usetikzlibrary{scopes}

\begin{document}

% \thispagestyle{empty}

\noindent {\bf MANUSCRIPT DN12236.hadj (PRD)}

\bigskip

\noindent {\bf Title: Scattering from compact objects: Regge poles and the complex angular momentum method}

\medskip
\noindent {\bf Authors: }\\
Mohamed Ould El Hadj \hspace{10pt} \href{https://orcid.org/0000-0002-8558-7992}{\textcolor{orcidlogocol}{ORCID :    }\textcolor{darkbue}{https://orcid.org/0000-0002-8558-7992}}\\
Tom Stratton \hspace{55pt}  \href{https://orcid.org/0000-0002-2936-5404}{\textcolor{orcidlogocol}{ORCID :    }\textcolor{darkbue}{https://orcid.org/0000-0002-2936-5404}}\\
Sam R. Dolan \hspace{53pt} \href{https://orcid.org/0000-0002-4672-6523}{\textcolor{orcidlogocol}{ORCID :    }\textcolor{darkbue}{https://orcid.org/0000-0002-4672-6523}}

\bigskip

\bigskip
\bigskip

\bigskip
\noindent {\bf Corrections corresponding to ``Specific Questions and
Comments to Address for This Paper''}

\bigskip
\noindent [A] - Item Q: All references include complete titles.

\bigskip
\noindent [B] - Item 1: WKB is an abbreviation in common use for more than half a century and accepted by PRD.

\bigskip
\noindent [C] - Item 2: Ok.

\bigskip
\noindent [D] - Item 3: Ok.

\bigskip
\noindent [E] - Item 4: Ok.

\bigskip
\bigskip

\bigskip
\noindent {\bf Other corrections in the text}

\bigskip
\noindent [I] - Line 725: ``\ldots CAM formulas to calculated scattering\ldots'' has to be changed into ``\ldots CAM formulas to calculate scattering \ldots''.

\bigskip\bigskip\bigskip\bigskip\bigskip

%\bigskip
%\noindent {\bf Important remark: We would like to check the new version before publication in PRD.}
%\bigskip

\noindent We thank you for your careful editing of our manuscript.

\noindent  M.O.E.H., T.S and S.R.D

\end{document}